---
title: "Join Lab"
author: "Quang Le"
date: "DATA 2401"
output: html_document
---


# Joins and `dplyr` practice

Load the `"nycflights13"` package.  You'll also need to load `dplyr` or `tidyverse`, if you didn't above.

```{r}
library(tidyverse)
#install.packages("nycflights13")
library(nycflights13)
my_flights <- flights # store my own version
my_flights <- na.omit(my_flights) ## this dataset has data on cancelled flights too, 
                                  ## so we 'll remove those for now

```


Create a dataframe of the average arrival delays for each _destination_

```{r}
average_arrival_delay <- my_flights %>% 
  group_by(dest) %>% 
  summarize(average_arrival_delay = mean(arr_delay,na.rm = TRUE)) %>% 
arrange(desc(average_arrival_delay))
head(average_arrival_delay)

```

Then use `left_join()` to join on the "airports" dataframe, which has the airport information.

```{r}
arrival_with_airports <- average_arrival_delay %>%
  left_join(airports, by = c("dest" = "faa"))
print(arrival_with_airports)

```


Note here: the column names don't match anymore! 

Which airport had the largest average arrival delay?

```{r}
largest_delay_airport <- arrival_with_airports %>%
  filter(!is.na(average_arrival_delay)) %>% 
  arrange(desc(average_arrival_delay)) %>%  
  slice(1) 
print(largest_delay_airport)
```



What is the average departure delay for each airline? To answer this with something more than just the FAA code of the airline name, you need to create a dataframe of the average departure delay for each _airline_, then use `left_join()` to join on the "airlines" dataframe.


```{r}
average_dep_delay <- my_flights %>% 
  group_by(carrier) %>% 
  summarize(average_dep_delay = mean(dep_delay,na.rm = TRUE)) %>% 
  arrange(desc(average_dep_delay))
head(average_dep_delay)

average_dep_airlines <- average_dep_delay %>% 
  left_join(airlines, by = "carrier")
head(average_dep_airlines)
```


Which airline had the smallest average departure delay? 



```{r}
smallest_delay_airlines <- average_dep_airlines %>%
  arrange(average_dep_delay) %>%  
  slice(1) 
print(smallest_delay_airlines)


```





Examples from R4DS:    

```{r}

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

```

Now you want to find each flight that went to one of those destinations. You could, of course, filter. But! Now you can semi-join: 

```{r}
flights %>% 
  semi_join(top_dest)

```


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?
Perform the two joins below and consider what the results are. 
```{r}
  anti_join(flights, airports, by = c("dest" = "faa"))
# anti_join keeps rows in the first dataframe (flights) where there is no match in the second dataframe (airports), based on the specified join condition.
# These destinations might represent missing or unrecorded airport codes in the airports dataset.
```
```{r}
  anti_join(airports, flights, by = c("faa" = "dest"))
# These are airports present in the airports dataset that do not have any flights recorded in the flights dataset.
```